---
title: "World Financial Indices"
author: "Victor Cruz Diaz, Brooks Lybrand, and Saeed Hashmi"
resource_files:
- .Renviron
- .Renviron
output:
  html_notebook:
    code_folding: hide
    toc: yes
    toc_depth: 4
    toc_float: yes
runtime: shiny
---

#Data.world Project 
This project explores several different financial indexes across 35 years for most countries in the world. The data and several insights can be found on [data.world](https://data.world/cruzdiaz-victor/s-18-edv-project-1).

#Load Packages
```{r setup, echo=FALSE}
library(tidyverse)
library(data.world)
library(DT)
library(plotly)
library(lubridate)
knitr::opts_chunk$set(echo = TRUE)
```

#Display Session Information
```{r}
sessionInfo()
```

#Retrieve countries from the financial indices dataset
The following query retrieves the distinct country names from the financial indices dataset at data.world and saves the country names in the country variable.
```{r}
project <- "https://data.world/cruzdiaz-victor/s18-edv-project-1-dataset"
data.world::set_config(cfg_env("DW_API"))
countries <- data.world::query(data.world::qry_sql(
 "   
	select distinct country_name
	from cleaned_data
 "),
dataset = project)
indices <- data.world::query(data.world::qry_sql(
 "   
	select distinct indicator_name
	from cleaned_data
 "),
dataset = project)
```

#Select which Data to Explore
Select the countries, financial indices, and years to query.

```{r}
inputPanel(
  selectInput("selectCountry_1", label = "Select Country",choices = countries, multiple=TRUE, selected=c("United States", "Mexico", "Canada")),
  selectInput("selectIndex_1", label = "Select Index",choices = indices, multiple=TRUE, selected=c("Financial Development Index", "Financial Institutions Access Index", "Financial Markets Index")),
  sliderInput("selectYear_1", label = "Year:", min = 1980, max = 2014, value = c(1980, 2014), step = 1)
)
```

#Create Dataframe with country_name, country_code, indicator_name, indicator_code, values, year data
Fetch the results from [data.world](https://data.world/cruzdiaz-victor/s18-edv-project-1-dataset) based on an SQL query created from the parameters selected above.

```{r}
df <- eventReactive(c(input$selectCountry_1, input$selectIndex_1, input$selectYear_1), {
  project <- "https://data.world/cruzdiaz-victor/s18-edv-project-1-dataset"
  data.world::set_config(cfg_env("DW_API"))
  paramQuery <- data.world::qry_sql(
    paste(
      "
  	    select country_name, country_code, indicator_name, indicator_code, values, year
        from cleaned_data c
        where cast(year as integer) between
      ", input$selectYear_1[1], "and", input$selectYear_1[2],
      "and country_name in (", paste("\"", paste(input$selectCountry_1, collapse="\", \""), "\"", sep=""), ")",
      "and indicator_name in (", paste("\"", paste(input$selectIndex_1, collapse="\", \""), "\"", sep=""), ")"
    )
   )
  data.world::query(paramQuery, dataset = project)
})
```
#Create a Countries Variable
Create a Countries variable with a string constructed from the selected regions in the input$selectCountry_1 variable.

```{r}
Countries <- eventReactive(c(input$selectCountry_1), {
  library('stringr')
  str_c(input$selectCountry_1, collapse=', ')
})
```

#Display the Dataframe as a Table
Here are the results of the query:

```{r}
# Notice the need to use df1() below:
renderDataTable({
  # DT::datatable(df() %>% dplyr::mutate(year = lubridate:year(order_date)), rownames = FALSE,
  DT::datatable(df(), rownames = FALSE, extensions = list(Responsive = TRUE, FixedHeader = TRUE))
})
```

#Render and Display the Years Boxplot
```{r}
tabsetPanel(
  tabPanel("Settings:"),# To create a tab panel - see https://shiny.rstudio.com/reference/shiny/latest/tabPanel.html
  tabPanel("Plot Size",
    numericInput("plotWidth_1", "Plot Width:", 800),
    numericInput("plotHeight_1", "Plot Height:", 600)),
  tabPanel("Plot Title",
    textInput("title_1", "Title", "Index Values"),
    numericInput("titleFont_1", "Title Font", 10)),
  tabPanel("Plot Legend",
    numericInput("legendTitleSize_1", "Legend Title Size", 10),
    numericInput("legendItemSize_1", "Legend Item Size", 10)),
  tabPanel("Axis Labels",
    textInput("xLabel_1", "x-Axis Label", "Years"),
    textInput("yLabel_1", "y-Axis Label", "Values"),
    numericInput("textFont_1", "textFont:", 10)),
  tabPanel("Data Size",
    numericInput("yDataMin_1", "yData Minimum (Required):", NA, min=0, step = 0.05),
    numericInput("yDataMax_1", "yData Maximum (Required):", NA, min=0, step = 0.05)))
```

```{r}
df1 <- eventReactive(c(input$selectCountry_1, input$selectIndex_1 ,input$selectYear_1, input$yDataMin_1, input$yDataMax_1), {
  if( ! is.na(input$yDataMin_1) & ! is.na(input$yDataMax_1)) {
    df() %>% dplyr::filter(between(values, input$yDataMin_1, input$yDataMax_1))
  }
  else {
    df()
  }
})
```

```{r}
# Boxplot - see http://ggplot2.tidyverse.org/reference/geom_boxplot.html
# Notice the need to use df(), and Countries() below:
renderPlotly({
  plot = df1() %>% ggplot() +
    geom_boxplot(mapping = aes(x = year, y=values, colour = country_name)) +

    theme(plot.title = element_text(size = input$titleFont_1, face = "bold")) +
    theme( # Legend Attributes - see https://github.com/tidyverse/ggplot2/wiki/Legend-Attributes
      legend.title=element_text(size=input$legendTitleSize_1),
      legend.text=element_text(size=input$legendItemSize_1)) +
    theme(axis.text=element_text(size=input$textFont_1),
          axis.title=element_text(size=input$textFont_1, face="bold"),
          axis.text.x = element_text(angle = 90, hjust = 1)) +
    theme(plot.margin=unit(c(2,1,1,1),"cm")) +
    scale_y_continuous(labels = scales::comma) + # Disable scientific notation
    scale_x_continuous(breaks=seq(1980, 2014, 2)) +
    ggtitle(paste(Countries(), input$title_1)) +
    xlab(input$xLabel_1) + ylab(input$yLabel_1) +
    theme_bw()
  ggplotly(plot, tooltip = c("values"), session="knitr", width = input$plotWidth_1, height = input$plotHeight_1)
})
```

<br/><br/><br/><br/><br/><br/><br/><br/><br/><br/>

#Render and Display Country Financial Indices Boxplots
```{r}
tabsetPanel(
  tabPanel("Settings:"),# To create a tab panel - see https://shiny.rstudio.com/reference/shiny/latest/tabPanel.html
  tabPanel("Plot Size",
    numericInput("plotWidth_2", "Plot Width:", 800),
    numericInput("plotHeight_2", "Plot Height:", 600)),
  tabPanel("Plot Title",
    textInput("title_2", "Title", "Index Values"),
    numericInput("titleFont_2", "Title Font", 10)),
  tabPanel("Plot Legend",
    numericInput("legendTitleSize_2", "Legend Title Size", 10),
    numericInput("legendItemSize_2", "Legend Item Size", 10)),
  tabPanel("Axis Labels",
    textInput("xLabel_2", "x-Axis Label", "Years"),
    textInput("yLabel_2", "y-Axis Label", "Values"),
    numericInput("textFont_2", "textFont:", 10)),
  tabPanel("Data Size", 
    numericInput("yDataMin_2", "yData Minimum (Required):", NA, min=0, step = 0.05),
    numericInput("yDataMax_2", "yData Maximum (Required):", NA, min=0, step = 0.05)))
```

```{r}
df2 <- eventReactive(c(input$selectCountry_2, input$selectIndex_2 ,input$selectYear_2, input$yDataMin_2, input$yDataMax_2), {
  if( ! is.na(input$yDataMin_2) & ! is.na(input$yDataMax_2)) {
    df() %>% dplyr::filter(between(values, input$yDataMin_2, input$yDataMax_2))
  }
  else {
    df()
  }
})
```
```{r}
# Boxplot - see http://ggplot2.tidyverse.org/reference/geom_boxplot.html
# Notice the need to use df(), and Countries() below:
renderPlotly({
  plot = df2() %>% ggplot() +
    geom_boxplot(mapping = aes(x = country_name, y = values, colour = country_name)) +

    theme(plot.title = element_text(size = input$titleFont_2, face = "bold")) +
    theme( # Legend Attributes - see https://github.com/tidyverse/ggplot2/wiki/Legend-Attributes
      legend.title=element_text(size=input$legendTitleSize_2),
      legend.text=element_text(size=input$legendItemSize_2)) +
    theme(axis.text=element_text(size=input$textFont_2),
          axis.title=element_text(size=input$textFont_2, face="bold"),
          axis.text.x = element_text(angle = 90, hjust = 1)) +
    theme(plot.margin=unit(c(2,1,1,1),"cm")) +
    scale_y_continuous(labels = scales::comma) + # Disable scientific notation
    ggtitle(paste(Countries(), input$title_2)) +
    xlab(input$xLabel_2) + ylab(input$yLabel_2) +
    theme_bw()
  ggplotly(plot, tooltip = c("values"), session="knitr", width = input$plotWidth_2, height = input$plotHeight_2)
})
```

<br/><br/><br/><br/><br/><br/><br/><br/><br/><br/>

# Insights
Here is the link to our [insights on data.world](https://data.world/cruzdiaz-victor/s-18-edv-project-1/insights).

#Summary of our insights regarding outliers in the data 

One thing that caught our attention about the data were the outliers, specifically those index values which were greater than 1. We did a bit of research to find out why they reached such abnormal level. One of our members found out that the Republic of Congo, Mali, Libya, and other nations experienced civil unrest which caused these abnormalities in the data. Almost every sharp spike or decrease in the values correlates with the state of the government of said country at that time.

Further information can be found on our [insights page](https://data.world/cruzdiaz-victor/s-18-edv-project-1/insights).

```{r}
get_outliers_df <- eventReactive(c(input$selectCountry_3, input$selectValues_3, input$selectYear_3), {
  project <- "https://data.world/cruzdiaz-victor/s18-edv-project-1-dataset"
  data.world::set_config(cfg_env("DW_API"))
  paramQuery <- data.world::qry_sql(
    paste(
      "
  	    select country_name, values, year, indicator_name
        from cleaned_data c
        where cast(year as integer) between
      ", input$selectYear_3[1], "and", input$selectYear_3[2],
      "and country_name in (", paste("\"", paste(input$selectCountry_3, collapse="\", \""), "\"", sep=""), ")"
      
    )
   )
  data.world::query(paramQuery, dataset = project)
})
```


```{r}
inputPanel(
      selectInput("selectCountry_3", label = "Select Country",choices = countries, multiple=TRUE, 
      selected=c("Sudan", "Libya", "Mali","Congo, Democratic Republic of")),

  sliderInput("selectYear_3", label = "Year:", min = 1980, max = 2014, value = c(2006,2014), step = 1)
)
```


```{r}
tabsetPanel(
  tabPanel("Settings:"),
  tabPanel("Plot Size",
    numericInput("plotWidth_3", "Plot Width:", 800),
    numericInput("plotHeight_3", "Plot Height:", 600)),
  tabPanel("Plot Title",
    textInput("title_3", "Title","2006 to 2014"),
    numericInput("titleFont_3", "Title Font", 20)),
  tabPanel("Axis Labels",
    textInput("xLabel_3", "x-Axis Label", "Countries"),
    textInput("yLabel_3", "y-Axis Label", "Values"),
    numericInput("textFont_3", "textFont:", 10)),
  tabPanel("Data Size", 
    numericInput("yDataMin_3", "yData Minimum (Required):", NA, min=0, step = 0),
    numericInput("yDataMax_3", "yData Maximum (Required):", NA, min=0, step = 0.55)))
```

```{r}
outliers_df <- eventReactive(c(input$selectCountry_3, input$selectValues_3 ,input$selectYear_3 ,input$yDataMin_3, input$yDataMax_3), {
  if( ! is.na(input$yDataMin_3) & ! is.na(input$yDataMax_3)) {
    get_outliers_df() %>% dplyr::filter(between(values, input$yDataMin_3, input$yDataMax_3))
  }
  else {
    get_outliers_df()
  }
})
```


```{r}

renderPlotly({
  plot = outliers_df() %>% ggplot() +
    geom_boxplot(mapping = aes(x = country_name, y = values, color = year )) +

    theme(plot.title = element_text(size = input$titleFont_3, face = "bold")) +
    theme(axis.text=element_text(size=input$textFont_3),
          axis.title=element_text(size=input$textFont_3, face="bold"),
          axis.text.x = element_text(angle = 150, hjust = 1)) +
    theme(plot.margin=unit(c(2,1,1,1),"cm")) +
    scale_y_continuous(labels = scales::comma) + 
    ggtitle(paste("Historic Outliers For The Years", input$title_3)) +
    xlab(input$xLabel_3) + ylab(input$yLabel_3) +
    theme_bw()
  ggplotly(plot, tooltip = c("values"), session="knitr", width = input$plotWidth_3, height = input$plotHeight_3)
})
```


<br/><br/><br/><br/><br/><br/><br/><br/><br/><br/>

#The Impact of Financial Crises

One of the interesting things that our team found was the impact of the 2008 recession on our data. The European continent, U.K, Australia, and the U.S experienced a decrease of various magnitudes in their financial indices due to the crisis, but China did not experience such a decrease in 2008. These changes could be attributed to the global nature of our financial stystems.

Further information can be found on our [insights page](https://data.world/cruzdiaz-victor/s-18-edv-project-1/insights).
```{r}
financial_crisis_df <- eventReactive(c(input$selectCountry_1, input$selectIndex_1, input$selectYear_1), {
  project <- "https://data.world/cruzdiaz-victor/s18-edv-project-1-dataset" 
  data.world::set_config(cfg_env("DW_API")) 
  paramQuery <- data.world::qry_sql(
    paste(
      "
  	    select country_name, country_code, indicator_name, indicator_code, values, year
        from cleaned_data c
        where cast(year as integer) between 2005 and 2011
        and (country_name='Australia'
        or country_name='United States'
        or country_name='China, P.R.: Mainland'
        or country_name='Europe'
        or country_name='United Kingdom')
      "
    )
   )
  data.world::query(paramQuery, dataset = project)
})
```

```{r}
renderPlotly({
  plot = financial_crisis_df() %>% ggplot() + 
    geom_boxplot(mapping = aes(x = year, y = values, fill = country_name)) +
    scale_y_continuous(labels = scales::comma) + # Disable scientific notation
    scale_x_continuous(breaks=seq(2004, 2012, 2)) +
    xlab("Years") + ylab("Values") +
    theme_bw() +
    facet_grid(~country_name)
  ggplotly(plot, tooltip = c("values"), width = 950, session="knitr")
})
```
